### 2021-02-18 ch06_paper_gpp/paper_gpp.tex
Added text to chapter 06

Due to complex feedback processes connected to the global carbon cycle
(concentration-carbon feedback and climate-carbon feedback; see
\cref{subsec:02:carbon_cycle_perturbations}), the response of the terrestrial
carbon cycle to changes in the atmospheric \ac{CO2} concentration and climate
is a major source of uncertainty in climate projections \autocite{Bodman2013,
  Booth2012, Collins2013}. This applies for example to multi-model projections
of future \ac{GPP} of the \acs{CMIP}5 ensemble. Thus, the aim of this study is
to reduce uncertainties in multi-model projections of future \ac{GPP} in the
\acs{CMIP}5 \acs{RCP}8.5 scenario at the end of the \nth{21} century with a new
\ac{ML}-based weighting technique and appropriate observations. The new
two-step approach utilizes aspects of emergent constraints and the \ac{MDER}
technique in combination with a \ac{ML} algorithm. In the first step, we apply
an existing emergent constraint on \ac{CO2} fertilization \autocite{Wenzel2016}
to constrain the \acp{ESM}' responses to rising atmospheric \ac{CO2}
concentration using observations of the increase of the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} \autocite{Keeling2005}. In a second step, we introduce a
supervised \ac{ML} algorithm based on boosting trees \autocite{Friedman2001} to
learn an empirical spatial relationship that links grid-wise future \ac{GPP} to
historical processes relevant to its simulation under present-day conditions.
In combination with observational products of the predictors, that relationship
can be used to further constrain uncertainties in the projected spatial maps of
\ac{GPP} at the end of the \nth{21} century in the \acs{RCP}8.5 scenario. We
examine both constraining the absolute \ac{GPP} and the fractional change in
\ac{GPP} as two independent approaches and target variables. Unlike univariate
linear regression used in the \ac{MDER} algorithm, the proposed \ac{GBRT}
algorithm is able to handle multiple predictors and copes with non-linearities
in the data. \ac{GBRT} is a well-known and successful tool used for
interpolation, classification and prediction in other fields of data science
and engineering \autocite{Death2007, Elith2008}. In the context of climate
science, \ac{GBRT} was recently applied to identify the key drivers of spatial
variations of the ratio of plant transpiration to total terrestrial
evapotranspiration in \acp{ESM} \autocite{Lian2018}.

@@

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{ch06_paper_gpp/figs/1.pdf}
  \caption{Schematic illustration of our two-step approach. In step 1 an
    emergent constraint by \textcite{Wenzel2016} is used to constrain the
    global mean fractional change in \acf{GPP} over the 21st century.
    Moreover, this constraint is used to rescale two different gridded target
    variables: absolute \acs{GPP} at the end of the \nth{21} century and
    fractional \acs{GPP} change over the \nth{21} century. In step 2, a
    \acl{ML} model is used to constrain these two target variables (step 2a:
    absolute \acs{GPP}; step 2b: fractional \acs{GPP} change) in two
    independent approaches. \AdaptedFrom{Schlund2020}. \cref{eq:06:xxx}}
  \label{fig:06:schematic_steps}
\end{figure}

@@

\begin{table}[t]
  \centering
  \begin{tabular}[t]{p{0.27\columnwidth} l p{0.11\columnwidth}
      p{0.3\columnwidth}}
    \toprule
    Predictor & \makecell{Observation‐driven \\ data} & Used time range &
    Physical connection to \acs{GPP} \\
    \midrule
    \predictor{GPP} & \makecell{FLUXNET-MTE \\ \autocite{Jung2011}} &
    \range{1991}{2000} & -- \\
    \midrule
    \predictor{LAI} & \makecell{LAI3g \\ \autocite{Zhu2013}} &
    \range{1982}{2005} & \acs{LAI} is a measure for the number of leaves in a
    grid cell. The photosynthesis rate is highly dependent on the number of
    leaves (and vegetation in general). \\
    \midrule
    \predictor{PR} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & Water is essential for the chemical processes of
    photosynthesis. \\
    \midrule
    \predictor{RSDS} & \makecell{ERA-Interim \\ \autocite{Dee2011}} &
    \range{1979}{2005} & Solar radiation is essential for the chemical
    processes of photosynthesis. \\
    \midrule
    \predictor{T} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & \acs{T} and photosynthesis rate have a common driver
    (incoming solar radiation). \\
    \bottomrule
  \end{tabular}
  \caption{Process-oriented diagnostics (\enquote{predictors} or
    \enquote{features}) used in the \acf{GBRT} model to predict the target
    variables. For step 2a (target variable: absolute \acs{GPP}), all listed
    variables are monthly climatologies of the specified time ranges in the
    historical climate. For step 2b (target variable: fractional \acs{GPP}
    change), the temporal mean over the specified time ranges is calculated for
    all variables. \AdaptedFrom{Schlund2020}.}
  \label{tab:06:predictors}
\end{table}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.37\columnwidth}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/3a.pdf}
    \caption{}
    \label{fig:06:step1:a}
  \end{subfigure}
  \begin{subfigure}[b]{0.37\columnwidth}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/3b.pdf}
    \caption{}
    \label{fig:06:step1:b}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\columnwidth}
    \raisebox{12.9mm}{\includegraphics[width=\columnwidth]{
      ch06_paper_gpp/figs/3c.pdf}}
  \end{subfigure}
  \caption{Emergent relationship between the global mean fractional change in
    \acf{GPP} and the sensitivity of the \acs{CO2} seasonal cycle amplitude to
    rising atmospheric \acs{CO2} concentrations observed at \acf{KUM}. Colored
    points refer to \acs{CMIP}5 models, the orange line and shaded area to the
    linear regression fit and its corresponding \acl{SPE} and the dashed black
    lines to the observational constraint. (a) Similar to
    \textcite{Wenzel2016}: The global mean fractional \acs{GPP} change after
    \acs{CO2} doubling in a \acf{1BGC} defined as 10-year mean of the
    \acs{1BGC} run centered at the time of \acs{CO2} doubling relative to the
    10-year mean of pre-industrial control conditions at the beginning of the
    \acs{1BGC} run. The sensitivity of the \acs{CO2} amplitude for the
    \acs{CMIP}5 models is calculated from the years \range{1860}{2005}, the
    corresponding observational value from the years \range{1979}{2019}. The
    constrained global mean \acs{GPP} change after \acs{CO2} doubling in the
    \acs{1BGC} run is $\pmrangeunit{30}{9}{\%}$. (b) Fractional global mean
    \acs{GPP} change over the \nth{21} century calculated from the 10-year
    mean \acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
    the emission-driven fully coupled \acs{RCP}8.5 simulations and the 10-year
    mean \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in
    the emission‐driven fully-coupled historical run (see \cref{eq:06:xxx}).
    In contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
    amplitude is calculated from the years \range{1979}{2019} for climate
    models and observations for better comparability (the historical
    \acs{CMIP}5 simulations are extended with the \acs{RCP}8.5 simulations for
    the years \range{2006}{2019}). The constrained global mean \acs{GPP}
    change over the \nth{21} century is $\pmrangeunit{39}{7}{\%}$.
    \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step1}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5a.pdf}
    \caption{}
    \label{fig:06:step2a_results:a}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5b.pdf}
    \caption{}
    \label{fig:06:step2a_results:b}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5c.pdf}
    \caption{}
    \label{fig:06:step2a_results:c}
  \end{subfigure}
  \caption{(a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the absolute
    \acf{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
    \acs{RCP}8.5 scenario using \cref{eq:06:xxx} (r\acs{MMM}) and its
    unweighted version (\acs{MMM}). The plot shows an almost constant value
    over the whole globe with a mean of $0.92$, which corresponds to the ratio
    of the constrained global mean \acs{GPP} change over the \nth{21} century
    from step 1 ($39 \unit{\%}$) and the \acs{CMIP}5 ensemble mean global mean
    \acs{GPP} change ($43 \unit{\%}$). All values close to $0$ for the dataset
    in the denominator have been masked to avoid divisions by $0$. (b) Bias
    between r\acs{MMM} and our \acf{GBRT} prediction for the end of the
    \nth{21} century. This corresponds to step 2a of our approach. (c) Bias
    between the modeled \acs{GPP} in the \acs{CMIP}5 \acs{MMM} of the
    historical simulation and the FLUXNET-MTE observation‐based estimate of
    \acs{GPP} \autocite{Jung2011} averaged between 1991 and 2000. Over large
    swaths of the globe, the \acs{CMIP}5 ensemble overestimates \acs{GPP} (red
    color). Panels (b) and (c) show similar bias patterns (pattern correlation
    of $R^2 = 0.88$). Thus, the \acs{GBRT} prediction in step 2a is able to
    correct the historical bias of the \acs{CMIP}5 ensemble relative to the
    FLUXNET‐MTE product. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2a_results}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8a.pdf}
    \caption{}
    \label{fig:06:step2b_results:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8b.pdf}
    \caption{}
    \label{fig:06:step2b_results:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8c.pdf}
    \caption{}
    \label{fig:06:step2b_results:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8d.pdf}
    \caption{}
    \label{fig:06:step2b_results:d}
  \end{subfigure}
  \caption{Fractional change in \acf{GPP} over the \nth{21} century (2100
    versus 2000) in the \acs{RCP}8.5 scenario for different statistical models
    (step 2b): (a) \acs{CMIP}5 \acf{MMM} of the fractional \acs{GPP} change,
    (b) rescaled \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:xxx}, (c)
    linear \acf{LASSO} model, and (d) \acf{GBRT} model. The geographical
    patterns from the different statistical models are very similar and show a
    higher \acs{GPP} increase in high latitudes and a lower \acs{GPP} closer
    to the equator. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2b_results}
\end{figure}

>>>>

Due to complex feedback processes connected to the global carbon cycle
(concentration-carbon feedback and climate-carbon feedback; see
\cref{subsec:02:carbon_cycle_perturbations}), the response of the terrestrial
carbon cycle to changes in the atmospheric \ac{CO2} concentration and climate
is a major source of uncertainty in climate projections \autocite{Bodman2013,
  Booth2012, Collins2013}. This applies for example to multi-model projections
of future \ac{GPP} of the \acs{CMIP}5 ensemble. Thus, the aim of this study is
to reduce uncertainties in multi-model projections of future \ac{GPP} in the
\acs{CMIP}5 \acs{RCP}8.5 scenario at the end of the \nth{21} century with a new
\ac{ML}-based weighting technique and appropriate observations. The new
two-step approach utilizes aspects of emergent constraints and the \ac{MDER}
technique in combination with a \ac{ML} algorithm. In the first step, we apply
an existing emergent constraint on \ac{CO2} fertilization \autocite{Wenzel2016}
to constrain the \acp{ESM}' responses to rising atmospheric \ac{CO2}
concentration using observations of the increase of the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} \autocite{Keeling2005}. In a second step, we introduce a
supervised \ac{ML} algorithm based on boosting trees \autocite{Friedman2001} to
learn an empirical spatial relationship that links grid-wise future \ac{GPP} to
historical processes relevant to its simulation under present-day conditions.
In combination with observational products of the predictors, that relationship
can be used to further constrain uncertainties in the projected spatial maps of
\ac{GPP} at the end of the \nth{21} century in the \acs{RCP}8.5 scenario. We
examine both constraining the absolute \ac{GPP} and the fractional change in
\ac{GPP} as two independent approaches and target variables. Unlike univariate
linear regression used in the \ac{MDER} algorithm, the proposed
\emph{\ac{GBRT}} algorithm is able to handle multiple predictors and copes with
non-linearities in the data. \ac{GBRT} is a well-known and successful tool used
for interpolation, classification and prediction in other fields of data
science and engineering \autocite{Death2007, Elith2008}. In the context of
climate science, \ac{GBRT} was recently applied to identify the key drivers of
spatial variations of the ratio of plant transpiration to total terrestrial
evapotranspiration in \acp{ESM} \autocite{Lian2018}.

@@

Here we briefly describe the data and methods used in this study.
\Cref{subsec:06:step_1} reviews the method of a previously published emergent
constraint on \ac{CO2} fertilization from \textcite{Wenzel2016} that we apply
here to our study to rescale the climate model output (step 1). This includes a
discussion of the caveats that are associated with this emergent constraint
that attributes changes in the observed increase in the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} entirely to the increase in atmospheric \ac{CO2}
concentration, yet other studies provide indications that part of this increase
is due to climate and land use changes \autocite{Bastos2019, Forkel2016,
  Piao2018, Zhao2016}. In \cref{subsec:06:step_2} the \ac{ML} technique used in
step 2 to constrain gridded \ac{GPP} projections is presented.
\Cref{fig:06:schematic_steps} shows an overview of our two-step approach.
\Cref{subsec:06:uncertainty_and_skill} describes how uncertainty and predictive
skill are quantified. A complete list of all \acs{CMIP}5 models used in this
study is given in \cref{tab:app:b:cmip5_models}. To apply the emergent
constraint in step 1, we need emission-driven \acs{CMIP}5 simulations. The
number of \acp{ESM} included is therefore seven, as in \textcite{Wenzel2016}.
More details on the methods and the experimental setup are given in
\cref{sec:app:si_for_paper_gpp}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{ch06_paper_gpp/figs/1.pdf}
  \caption{Schematic illustration of our two-step approach. In step 1 an
    emergent constraint by \textcite{Wenzel2016} is used to constrain the
    global mean fractional change in \acf{GPP} over the 21st century.
    Moreover, this constraint is used to rescale two different gridded target
    variables: absolute \acs{GPP} at the end of the \nth{21} century and
    fractional \acs{GPP} change over the \nth{21} century. In step 2, a
    \acl{ML} model is used to constrain these two target variables (step 2a:
    absolute \acs{GPP}; step 2b: fractional \acs{GPP} change) in two
    independent approaches. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:schematic_steps}
\end{figure}

In this section, we present the observational emergent constraint on the
\ac{CO2} fertilization effect by \textcite{Wenzel2016} which we use to globally
rescale the models for their individual bias in the sensitivity of \ac{GPP} to
an increase in atmospheric \ac{CO2} concentration. Our objective here is to
present a general methodology and suggest further improvements upon this
constraint that could be used in the future. \textcite{Wenzel2016} show a
correlation between the increase in the \ac{CO2} seasonal cycle amplitude
simulated by models and the projected \ac{CO2} fertilization. When combined
with the observed trends in the \ac{CO2} seasonal cycle amplitude, the
fractional \ac{GPP} change at the time of \ac{CO2} doubling relative to
pre-industrial conditions in a \ac{1BGC} is constrained. The predictor in this
study is the sensitivity of the \ac{CO2} seasonal cycle amplitude to rising
atmospheric \ac{CO2} concentrations, defined as the slope of the linear
regression between the \ac{CO2} seasonal cycle amplitude and the annual mean
atmospheric \ac{CO2} concentrations (see \cref{fig:app:b:co2} for details). The
emergent constraint is physically motivated by the hypothesis that increasing
terrestrial \ac{GPP} is a main driver for the observed changes in the \ac{CO2}
seasonal cycle amplitude \autocite{Gray2014, Keeling1996, Zhao2014}. This
hypothesis is based on the fact that the seasonal cycle in atmospheric \ac{CO2}
concentration originates from photosynthesis and decomposition processes: in
summer, photosynthesis removes \ac{CO2} from the atmosphere; in winter,
additional \ac{CO2} is added to the atmosphere due to the decomposition of
organic matter \autocite{Keeling1995}. Thus, \ac{CO2} fertilization leads to an
increase of the \acs{CO2} seasonal cycle amplitude due to increased \acs{CO2}
removal from the atmosphere. We note that this emergent constraint implicitly
assumes that there is no temperature-driven increase in respiration over time.
Therefore, this approach most likely overestimates the true sensitivity of
\ac{GPP} to an increase in the atmospheric \ac{CO2} concentration.

However, we note that this emergent constraint is not undisputed. In other
studies, the observed change in the \ac{CO2} seasonal cycle amplitude is
attributed to other factors. Using a dynamic global vegetation model,
\textcite{Forkel2016} find that the increase in the \ac{CO2} seasonal cycle
amplitude above $40 \unit{\degree N}$ is mainly driven by the response of
plants to global warming instead of \ac{CO2}. Similarly, \textcite{Piao2018}
show that elevated atmospheric \acs{CO2} concentration and climate change
equally contribute to the increase in the \ac{CO2} seasonal cycle in northern
high latitudes with additional smaller contributions from air-sea carbon fluxes
and land use changes. For their analysis, \textcite{Piao2018} use \ac{CO2}
records from 26 stations and nine terrestrial ecosystem models. An evaluation
of two atmospheric inversions and eleven land surface models shows contrasting
effects of \ac{CO2} fertilization (positive) and global warming (negative) on
the \ac{CO2} seasonal cycle amplitude increase in northern high latitudes with
negligible contributions from land use changes \autocite{Bastos2019}. A further
study by \textcite{Zhao2016} finds that for seven out of nine dynamic
vegetation models, the \ac{CO2} fertilization effect is the strongest driver
for the observed changes in the \ac{CO2} seasonal cycle amplitude, while the
two remaining models equally attribute this to \ac{CO2} fertilization, climate
change and land use changes. Moreover, \textcite{Winkler2019a} discuss
limitations of carbon cycle emergent constraints related to the calculation of
the predictor from observation-based data and the inter-model relationship. As
a case study, they adopt a further emergent constraint by
\textcite{Winkler2019} that uses the response of the \ac{LAI} to ambient
\ac{CO2} instead of the \ac{CO2} seasonal cycle amplitude sensitivity to
constrain the \ac{GPP} increase in the northern high latitudes. Furthermore, it
has been shown that the \acs{CMIP}5 models \autocite{Graven2013} and the TRENDY
dynamic vegetation models \autocite{Thomas2016} greatly underestimate long-term
changes in the \ac{CO2} seasonal cycle which questions their ability to
attribute changes in the \ac{CO2} seasonal cycle to different contributors.
Nevertheless, we emphasize that we here present a generic methodology and use
the results of an existing emergent constraint. Our proposed framework is
universal and could be based on other emergent constraints in the future.

In contrast to \textcite{Wenzel2016}, instead of idealized simulations we apply
the presented emergent constraint on the emission-driven \acs{RCP}8.5
simulation \autocite{Riahi2011} and examine whether it still holds for this
simulation where forcings other than \ac{CO2} also change. Moreover, we
consider the global mean \ac{GPP} change instead of the \ac{GPP} change in the
northern extratropics as in \textcite{Wenzel2016}. Due to the strong annual
mixing of \ac{CO2} in the atmosphere, the change in the \ac{CO2} seasonal cycle
amplitude at \ac{KUM} is representative for the whole globe. Thus, we apply
this emergent constraint to the global mean \ac{GPP} change. However, we note
that this approach might introduce errors since the \ac{CO2} fertilization
effect could be different for tropical and extratropical ecosystems due to
nutrient limitations and the temperature dependence of Rubisco kinetics
\autocite{CraftsBrandner2000} in the tropics. Nevertheless, we argue that we
can use the constraint if the emergent relationship still holds within our
considered framework (the \acs{RCP}8.5 scenario simulated by the \acs{CMIP}5
models).

Therefore, our target variable in step 1 is the global mean fractional
\acs{GPP} change over the \nth{21} century $f$ in the \acs{RCP}8.5 scenario,
which is calculated from the 10-year mean \ac{GPP} at the end of the \nth{21}
century in the emission‐driven fully-coupled \acs{RCP}8.5 simulations
(\range{2091}{2100}) and the 10-year mean \ac{GPP} at the end of the \nth{20}
century (\range{1991}{2000}) in the emission-driven historical simulation:
\begin{equation}
  f = \frac{\mathtxt{GPP (\range{2091}{2100})}}{
    \mathtxt{GPP (\range{1991}{2000})}} - 1.
  \label{eq:06:f}
\end{equation}
This variable is given as a percentage, where negative values correspond to a
decrease in \ac{GPP}, positive values to an increase in \ac{GPP} and a value of
$0 \unit{\%}$ to no \ac{GPP} change. The \ac{CO2} amplitude sensitivity is
calculated for the years \range{1979}{2019}, which is the full available time
range for the observed \ac{CO2} record at \ac{KUM} \autocite{Keeling2005}.
Similar \textcite{Wenzel2016}, we extract the observed \ac{CO2} amplitude from
smoothed atmospheric \ac{CO2} concentrations (using a stiff cubic spline
function plus a four-harmonics functions with linear gain) as provided by
\textcite{Keeling2005}. For the \acs{CMIP}5 models, the respective grid cell
closest to \ac{KUM} ($19.5 \unit{\degree N}$, $154.8 \unit{\degree W}$) is
extracted. Additionally, the emission-driven historical simulations are
extended with the emission-driven \acs{RCP}8.5 simulations for the years 2006
to 2019. More details including an illustration of this calculation are shown
in \cref{fig:app:b:co2}.

Since the global constraint above cannot be used directly as a predictor in the
\ac{ML} approach which operates on gridded values (see
\cref{subsec:06:step_2}), we use it as first step to globally rescale the
\ac{ESM} output. This important first step allows us to constrain the large
uncertainty in the global \ac{CO2} fertilization effect of the \acs{CMIP}5
models. In this study, two different gridded target variables are examined (see
\cref{fig:06:schematic_steps}): the monthly climatology of future absolute
\ac{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
\acs{RCP}8.5 scenario (see step 2a) and the fractional \acs{GPP} change over
the \nth{21} century as defined in \cref{eq:06:f} (see step 2b). The global
rescaling of the gridded target variables works as follows: Let $f_m$ be the
global mean fractional \acs{GPP} change over the \nth{21} century of \ac{ESM}
$m$, $f'$ the constrained value of the global mean fractional \ac{GPP} change
and $y_{mi}$ the target variable of climate model $m$. Then, we define the
rescaled target variable $y'_{mi}$ by
\begin{equation}
  y'_{mi} = y_{mi} \cdot \frac{f'}{f_m},
  \label{eq:06:y}
\end{equation}
where the index $i$ counts over all grid cells (and months when absolute
\ac{GPP} is used as target variable in step 2a) and $m$ refers to the \ac{ESM}
($m \in \left\{ 1, 2, \ldots, 7 \right\}$). Thus, values for the individual
grid cells/months are rescaled by a factor that is determined by the ratio
between the models' global mean \ac{GPP} change and the constrained global mean
\ac{GPP} change derived from the \textcite{Wenzel2016} emergent constraint
using the observed change in the \ac{CO2} cycle amplitude. Models that simulate
a too high global mean \ac{GPP} change are nudged toward smaller values and
vice versa. The rescaled target variables are then fed into the \ac{ML} model
(step 2) presented in the following section. We note that this global rescaling
assumes that the relative biases in the fractional \ac{GPP} change of the
individual \acp{ESM} are constant over the whole globe.

@@

\begin{table}[t]
  \centering
  \begin{tabular}[t]{p{0.27\columnwidth} l p{0.11\columnwidth}
      p{0.3\columnwidth}}
    \toprule
    Predictor & \makecell{Observation‐driven \\ data} & Used time range &
    Physical connection to \acs{GPP} \\
    \midrule
    \predictor{GPP} & \makecell{FLUXNET-MTE \\ \autocite{Jung2011}} &
    \range{1991}{2000} & -- \\
    \midrule
    \predictor{LAI} & \makecell{LAI3g \\ \autocite{Zhu2013}} &
    \range{1982}{2005} & \acs{LAI} is a measure for the number of leaves in a
    grid cell. The photosynthesis rate is highly dependent on the number of
    leaves (and vegetation in general). \\
    \midrule
    \predictor{PR} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & Water is essential for the chemical processes of
    photosynthesis. \\
    \midrule
    \predictor{RSDS} & \makecell{ERA-Interim \\ \autocite{Dee2011}} &
    \range{1979}{2005} & Solar radiation is essential for the chemical
    processes of photosynthesis. \\
    \midrule
    \predictor{T} & \makecell{CRU \\ \autocite{Harris2014}} &
    \range{1901}{2005} & \acs{T} and photosynthesis rate have a common driver
    (incoming solar radiation). \\
    \bottomrule
  \end{tabular}
  \caption{Process-oriented diagnostics (\emph{predictors} or \emph{features})
    used in the \acf{GBRT} model to predict the target variables. For step 2a
    (target variable: absolute \acs{GPP}), all listed variables are monthly
    climatologies of the specified time ranges in the historical climate. For
    step 2b (target variable: fractional \acs{GPP} change), the temporal mean
    over the specified time ranges is calculated for all variables.
    \AdaptedFrom{Schlund2020}.}
  \label{tab:06:predictors}
\end{table}

\subsection{Gradient Boosted Regression Tree (\acs{GBRT}) Algorithm (Step 2)}
\label{subsec:06:step_2}

For step 2 we use the \ac{GBRT} algorithm to find an optimal regression
function $F$ relating the target variable $y$ to a set of $K$ predictors
$\bm{x} = \left( x^{(1)}, x^{(2)}, \ldots, x^{(K)} \right)$ (\emph{features} or
\emph{covariates}) such that the predicted target variable is given by $\hat{y}
= F(\bm{x})$ \autocite{Death2007, Elith2008, Friedman2001, Friedman2002}. Its
basic elements are decision trees, which use binary splits of the input data to
create decision rules. Due to their simple nature, decision trees are easy to
interpret but cannot be used to create satisfying predictions for very complex
datasets. This issue can be overcome by a technique called \emph{boosting}
\autocite{Freund1996}. Boosting iteratively improves the performance of
\emph{weak learners} (in our case decision trees) in the following way: At
first, a single decision tree is fitted to a randomly selected subsample of the
training data $\left\{ \left( \bm{x}_i, y_i \right) \mid i \in I \right\}$ with
an arbitrary index set $I$ (using a predefined loss function, \eg{} the mean
square error). Then, further decision trees are linearly added to the
regression function $F$ (scaled by a regularization parameter $\nu \ll 1$
called \emph{learning rate}) and the model is fitted again in the same manner.
This ensures that the prediction for poorly modeled training points in the
beginning will gradually improve during the training. The whole procedure is
repeated until the loss function does not improve anymore, as evaluated on an
independent validation data set. Relevant hyperparameters of this algorithm
(maximum number of decision trees, maximum depth of the decision trees,
learning rate and the subsample ratio) are optimized by $5$-fold \ac{CV} using
the training data. More details on the \ac{GBRT} algorithm are given in
\cref{subsec:app:b:gbrt}.

\subsection{Assessment of Uncertainty and Predictive Skill}
\label{subsec:06:uncertainty_and_skill}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.37\columnwidth}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/3a.pdf}
    \caption{}
    \label{fig:06:step1:a}
  \end{subfigure}
  \begin{subfigure}[b]{0.37\columnwidth}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/3b.pdf}
    \caption{}
    \label{fig:06:step1:b}
  \end{subfigure}
  \begin{subfigure}[b]{0.24\columnwidth}
    \raisebox{12.9mm}{\includegraphics[width=\columnwidth]{
      ch06_paper_gpp/figs/3c.pdf}}
  \end{subfigure}
  \caption{Emergent relationship between the global mean fractional change in
    \acf{GPP} and the sensitivity of the \acs{CO2} seasonal cycle amplitude to
    rising atmospheric \acs{CO2} concentrations observed at \acf{KUM}. Colored
    points refer to \acs{CMIP}5 models, the orange line and shaded area to the
    linear regression fit and its corresponding \acl{SPE} and the dashed black
    lines to the observational constraint. (a) Similar to
    \textcite{Wenzel2016}: The global mean fractional \acs{GPP} change after
    \acs{CO2} doubling in a \acf{1BGC} defined as 10-year mean of the
    \acs{1BGC} run centered at the time of \acs{CO2} doubling relative to the
    10-year mean of pre-industrial control conditions at the beginning of the
    \acs{1BGC} run. The sensitivity of the \acs{CO2} amplitude for the
    \acs{CMIP}5 models is calculated from the years \range{1860}{2005}, the
    corresponding observational value from the years \range{1979}{2019}. The
    constrained global mean \acs{GPP} change after \acs{CO2} doubling in the
    \acs{1BGC} run is $\pmrangeunit{30}{9}{\%}$. (b) Fractional global mean
    \acs{GPP} change over the \nth{21} century calculated from the 10-year
    mean \acs{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in
    the emission-driven fully coupled \acs{RCP}8.5 simulations and the 10-year
    mean \acs{GPP} at the end of the \nth{20} century (\range{1991}{2000}) in
    the emission‐driven fully-coupled historical run (see \cref{eq:06:f}). In
    contrast to \textcite{Wenzel2016}, the sensitivity of the \acs{CO2}
    amplitude is calculated from the years \range{1979}{2019} for climate
    models and observations for better comparability (the historical
    \acs{CMIP}5 simulations are extended with the \acs{RCP}8.5 simulations for
    the years \range{2006}{2019}). The constrained global mean \acs{GPP}
    change over the \nth{21} century is $\pmrangeunit{39}{7}{\%}$.
    \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step1}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5a.pdf}
    \caption{}
    \label{fig:06:step2a_results:a}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5b.pdf}
    \caption{}
    \label{fig:06:step2a_results:b}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/5c.pdf}
    \caption{}
    \label{fig:06:step2a_results:c}
  \end{subfigure}
  \caption{(a) Ratio of the rescaled \acs{CMIP}5 ensemble mean of the absolute
    \acf{GPP} at the end of the \nth{21} century (\range{2091}{2100}) in the
    \acs{RCP}8.5 scenario using \cref{eq:06:y} (r\acs{MMM}) and its unweighted
    version (\acs{MMM}). The plot shows an almost constant value over the
    whole globe with a mean of $0.92$, which corresponds to the ratio of the
    constrained global mean \acs{GPP} change over the \nth{21} century from
    step 1 ($39 \unit{\%}$) and the \acs{CMIP}5 ensemble mean global mean
    \acs{GPP} change ($43 \unit{\%}$). All values close to $0$ for the dataset
    in the denominator have been masked to avoid divisions by $0$. (b) Bias
    between r\acs{MMM} and our \acf{GBRT} prediction for the end of the
    \nth{21} century. This corresponds to step 2a of our approach. (c) Bias
    between the modeled \acs{GPP} in the \acs{CMIP}5 \acs{MMM} of the
    historical simulation and the FLUXNET-MTE observation‐based estimate of
    \acs{GPP} \autocite{Jung2011} averaged between 1991 and 2000. Over large
    swaths of the globe, the \acs{CMIP}5 ensemble overestimates \acs{GPP} (red
    color). Panels (b) and (c) show similar bias patterns (pattern correlation
    of $R^2 = 0.88$). Thus, the \acs{GBRT} prediction in step 2a is able to
    correct the historical bias of the \acs{CMIP}5 ensemble relative to the
    FLUXNET‐MTE product. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2a_results}
\end{figure}

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8a.pdf}
    \caption{}
    \label{fig:06:step2b_results:a}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8b.pdf}
    \caption{}
    \label{fig:06:step2b_results:b}
  \end{subfigure}
  \\
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8c.pdf}
    \caption{}
    \label{fig:06:step2b_results:c}
  \end{subfigure}
  ~
  \begin{subfigure}[b]{\SubfigureWidth{}}
    \includegraphics[width=\columnwidth]{ch06_paper_gpp/figs/8d.pdf}
    \caption{}
    \label{fig:06:step2b_results:d}
  \end{subfigure}
  \caption{Fractional change in \acf{GPP} over the \nth{21} century (2100
    versus 2000) in the \acs{RCP}8.5 scenario for different statistical models
    (step 2b): (a) \acs{CMIP}5 \acf{MMM} of the fractional \acs{GPP} change,
    (b) rescaled \acs{CMIP}5 \acl{MMM} (r\acs{MMM}) using \cref{eq:06:y}, (c)
    linear \acf{LASSO} model, and (d) \acf{GBRT} model. The geographical
    patterns from the different statistical models are very similar and show a
    higher \acs{GPP} increase in high latitudes and a lower \acs{GPP} closer
    to the equator. \AdaptedFrom{Schlund2020}.}
  \label{fig:06:step2b_results}
\end{figure}
